---
title: "HW_01"
author: "Henry Traynor"
date: "Due: 25 August"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 60),
  echo = TRUE)
```

### Assignment

All code can be found here:
https://github.com/HenryTraynor/WILD8390/blob/main/Homework/HW_01.Rmd

Create a self-contained R script (or better yet an Rmarkdown file) to do the following:

Simulate logistic regression data according to $y_i \sim \textrm{Bern}(p_i)$ and $\textrm{logit}(p_i) = \beta_0 + \beta_1x_i$ with $\beta_0=-1$ and $\beta_1=1$. Generate the covariate using `x <- rnorm(n=100)`.

```{r}
# Set seed so results are the same in development and written answers
set.seed(314159)

# initialize parameters
beta_0 <- -1
beta_1 <- 1
# generate predictor variable
x <- rnorm(n=100)
# generate response
# 'plogis()' gives inverse 'logit()' to yield probability
y <- rbinom(100, 
            size=1, 
            prob = plogis(beta_0 + beta_1 * x))
```

Fit a logistic regression model to the simulated data using the `glm` function. Create a figure showing $x$ and $y$ with the fitted regression line. Use `predict` to get the fitted line.

```{r}
library(ggplot2)

logistic_regression <- glm(y ~ x,
                           family = binomial)
# df to store regression data
predicted_data <- data.frame(
  x = seq(min(x), max(x), length.out = 100)
)
predicted_data$y_hat <- predict(logistic_regression,
                                newdata = predicted_data,
                                type = "response")

# df to store model input data for plotting
data <- data.frame(x, y)

ggplot() +
  geom_point(
    data = data,
    aes(x = x, y = y, color = "Observed data")
  ) +
  geom_line(
    data = predicted_data,
    aes(x = x, y = y_hat, color = "Logistic regression")
  ) +
  labs(
    x = "x",
    y = "y",
    color = "Data"
  )
```

Fit the logistic regression model in JAGS. Use `Normal(0,var=10)` priors for the regression coefficients.

```{r}
library(rjags)
# Model created using reference of the JAGS Version 4.3.0 user manual and rjags documentation on rdocumentation.com . I will reference what section my code is coming from in comments.
# link: https://pilotfiber.dl.sourceforge.net/project/mcmc-jags/Manuals/4.x/jags_user_manual.pdf?viasf=1&fid=d949167c0f4eb392&e=1787674022&st=a8q7euJf-PwAbqj3yzUU_Q
jags_data <- list(
  y = y,
  x = x,
  n = length(y)
)

#Chapter 2: model creation
#Note that dnorm() in JAGS/BUGS takes precision, not variance, so I have used dnorm(0,1/10) for my priors.

string <- "model {
  for(i in 1:n) {
    logit(p[i]) <- beta_0 + beta_1 * x[i]
    y[i] ~ dbern(p[i])
  }
  beta_0 ~ dnorm(0, 1/10)
  beta_1 ~ dnorm(0, 1/10)
}
"

# https://www.rdocumentation.org/packages/rjags/versions/4-17/topics/jags.model
# I was unsure what satisfactory 'n.chains' or 'n.adapt' would be so I used values similar to those shown in documentation
jags_model <- jags.model(
  textConnection(string),
  data = jags_data,
  n.chains = 3,
  n.adapt = 1000
)

# burn-in period as described in 4.1 of JAGS manual
update(
  jags_model,
  n.iter = 1000
)

# posterior samples: 4.1 of JAGS manual
samples <- coda.samples(
  model = jags_model,
  variable.names = c("beta_0", "beta_1"),
  n.iter = 5000
)

summary(samples)
```

Compare and interpret the estimates of $\beta_0$ and $\beta_1$ from the classical analysis and the Bayesian analysis.

```{r}
# 'samples' is a mcmc.list so it needs to be a dataframe/matrix to pull values from
samples_matrix <- as.matrix(samples)

results <- data.frame(
  Parameter = c("beta_0", "beta_1"),
  GLM = coef(logistic_regression),
  JAGS = c(
    mean(samples_matrix[, "beta_0"]),
    mean(samples_matrix[, "beta_1"])
  ),
  JAGS_Lower = c(
    quantile(samples_matrix[, "beta_0"], 0.025),
    quantile(samples_matrix[, "beta_1"], 0.025)
  ),
  JAGS_Upper = c(
    quantile(samples_matrix[, "beta_0"], 0.975),
    quantile(samples_matrix[, "beta_1"], 0.975)
  )
)

# round our decimals for cleanliness' sake
results[, -1] <- round(results[, -1], 3)
results
```

Our GLM does a good job of estimating $\beta$ values. The Bayesian estimate from JAGS is less accurate but still captures the true values in its confidence intervals. Estimates for $\beta_0$ and $\beta_1$ differ by 0.006 and 0.021, respectively. Most importantly, both models correctly determined the sign of $\beta_1$, showing increases in $x$ are associated with an increasing probability of $y=1$.

I was curious to see the JAGS output plotted. So here is a figure showing both regressions together. Gray region represents the uncertainty provided by the Bayesian model.

```{r}
jags_predicted <- data.frame(
  x = seq(min(x), max(x), length.out = 100)
)

jags_predictions <- sapply(
  jags_predicted$x,
  function(x_value) {
    plogis(
      samples_matrix[, "beta_0"] +
      samples_matrix[, "beta_1"] * x_value
    )
  }
)

jags_predicted$mean <- apply(
  jags_predictions,
  2,
  mean
)
jags_predicted$lower <- apply(
  jags_predictions,
  2,
  quantile,
  0.025
)
jags_predicted$upper <- apply(
  jags_predictions,
  2,
  quantile,
  0.975
)

ggplot() +
  geom_point(
    data = data,
    aes(x = x, y = y, color = "Observed data")
  ) +
  geom_line(
    data = predicted_data,
    aes(x = x, y = y_hat, color = "GLM")
  ) +
  geom_ribbon(
    data = jags_predicted,
    aes(
      x = x,
      ymin = lower,
      ymax = upper
    ),
    alpha = 0.2
  ) +
  geom_line(
    data = jags_predicted,
    aes(x = x, y = mean, color = "JAGS")
  ) +
  labs(
    x = "x",
    y = "Probability of y = 1",
    color = "Model"
  ) +
  theme_minimal()
```