---
title: "HW_04"
author: "Henry Traynor"
date: 'Due: 5:00pm, 15 September'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a self-contained R script or Rmarkdown/Quarto file to do the following:

Using the simulated data, compare the prior and posterior predictive distributions of $\lambda$ for each of the 3 forest types. 

Hint: because forest type is categorical (not continuous), you can visualize the prior and posterior predictive distributions as histograms for each forest type.

```{r Data Generation}
set.seed(314159) # for reproducibility

nSites <- 100
nVisits <- 4

forest <- factor(sample(c("Hardwood", "Mixed", "Pine"), nSites, replace=TRUE))
forestMixed <- ifelse(forest=="Mixed", 1, 0)        ## Dummy
forestPine <- ifelse(forest=="Pine", 1, 0)          ## Dummy
temp <- matrix(rnorm(nSites*nVisits), nrow=nSites, ncol=nVisits)

# true coeffs
beta0 <- 0
beta1 <- -1
beta2 <- 1
lambda <- exp(beta0 + beta1*forestMixed + beta2*forestPine)
alpha0 <- -2
alpha1 <- 1
p <- plogis(alpha0 + alpha1*temp)

# simulate occupancy and detection data
N <- rpois(nSites, lambda=lambda)         ## local abundance 
y <- matrix(NA, nrow=nSites, ncol=nVisits)
for(i in 1:nSites) {
    y[i,] <- rbinom(nVisits, size=N[i], prob=p[i,])
}

# summary statistics
maxCounts <- apply(y, 1, max) 
table(maxCounts)            

naiveOccupancy <- sum(maxCounts>0)/nSites
naiveOccupancy 

naiveAbund <- sum(maxCounts)
naiveAbund
```

```{r Bayesian Modeling}
# prep data in named list
jags.data <- list(y=y,
                  forestMixed=forestMixed,
                  forestPine=forestPine,
                  nSites=nSites, 
                  temp=temp,
                  nOccasions=nVisits)
# initial values for MCMC
jags.inits <- function() {
    list(beta0=rnorm(1), alpha0=rnorm(1), N=maxCounts)
}
# parameters to save
jags.pars <- c("beta0", "beta1", "beta2",
               "alpha0", "alpha1", "totalAbundance")

# run model
library(jagsUI)
jags.post.samples <- jags.basic(data=jags.data, 
                                inits=jags.inits,
                                parameters.to.save=c(jags.pars,"N"), 
                                model.file="Nmix-model-covs.jag",
                                n.chains=3,
                                n.adapt=100, 
                                n.burnin=0,
                                n.iter=2000,
                                parallel=TRUE)
```

```{r Summary & Plots}
summary(jags.post.samples[,jags.pars])
plot(jags.post.samples[,paste0("N[", 1:4, "]")])
plot(jags.post.samples[,jags.pars[1:3]])
plot(jags.post.samples[,jags.pars[4:5]])
```

```{r Bayesian Prediction}
p.coef.post <- as.matrix(jags.post.samples[,c("alpha0","alpha1")])
head(p.coef.post, n=4)

n.iter <- nrow(p.coef.post)
temp.pred <- seq(-3, 3, length=50)
p.post.pred <- matrix(NA, nrow=n.iter, ncol=length(temp.pred))

for(i in 1:n.iter) {
    p.post.pred[i,] <- plogis(p.coef.post[i,"alpha0"] +
                              p.coef.post[i,"alpha1"]*temp.pred)
}

plot(temp.pred, p.post.pred[1,], type="l", xlab="Temperature (standardized)",
     ylab="Detection probability", ylim=c(0, 1), col=gray(0.8))
for(i in seq(1, n.iter, by=10)) {  ## Thin by 10
    lines(temp.pred, p.post.pred[i,], col=gray(0.8))  }
pred.post.mean <- colMeans(p.post.pred)
pred.post.lower <- apply(p.post.pred, 2, quantile, prob=0.025)
pred.post.upper <- apply(p.post.pred, 2, quantile, prob=0.975)
lines(temp.pred, pred.post.mean, col="blue")
lines(temp.pred, pred.post.lower, col="blue", lty=2)
lines(temp.pred, pred.post.upper, col="blue", lty=2)
```

```{r Bayesian Prediction}
lambda.coef.post <- as.matrix(jags.post.samples[,c("beta0","beta1", "beta2")])
head(lambda.coef.post, n=4)

n.iter <- nrow(lambda.coef.post)
forest.pred <- c(rep("Hardwood", 50),
                 rep("Mixed", 50),
                 rep("Pine",50)
                 )

forest.Mixed.pred <- ifelse(forest.pred=="Mixed", 1, 0)
forest.Pine.pred <- ifelse(forest.pred=="Pine", 1, 0)

lambda.post.pred <- matrix(NA, nrow=n.iter, ncol=length(forest.pred))

for(i in 1:n.iter) {
    lambda.post.pred[i,] <- plogis(lambda.coef.post[i,"beta0"] + lambda.coef.post[i,"beta1"]*forest.Mixed.pred + lambda.coef.post[i, "beta2"]*forest.Pine.pred)
}

#-------------
# left here; rest needs to be turned into forest & lambda
#-------------
for(i in seq(1, n.iter, by=10)) {  ## Thin by 10
    lines(temp.pred, p.post.pred[i,], col=gray(0.8))  }
pred.post.mean <- colMeans(p.post.pred)
pred.post.lower <- apply(p.post.pred, 2, quantile, prob=0.025)
pred.post.upper <- apply(p.post.pred, 2, quantile, prob=0.975)
lines(temp.pred, pred.post.mean, col="blue")
lines(temp.pred, pred.post.lower, col="blue", lty=2)
lines(temp.pred, pred.post.upper, col="blue", lty=2)
```

```{r Prior Sampling}
jags.data2 <- jags.data
jags.data2$y[] <- NA     # Replace data with missing values

jags.prior.samples <- jags.basic(data=jags.data2,
                                 inits=jags.inits,
                                 parameters.to.save=jags.pars,
                                 model.file="Nmix-model-covs.jag",
                                 n.chains=3,
                                 n.adapt=100,
                                 n.burnin=0,
                                 n.iter=2000,
                                 parallel=TRUE)

summary(jags.prior.samples)
```

```{r Prior Prediction}
p.coef.prior <- as.matrix(jags.prior.samples[,c("alpha0","alpha1")])
p.prior.pred <- matrix(NA, nrow=n.iter, ncol=length(temp.pred))
for(i in 1:n.iter) {
    p.prior.pred[i,] <- plogis(p.coef.prior[i,"alpha0"] +
                               p.coef.prior[i,"alpha1"]*temp.pred)
}

pred.prior.mean <- colMeans(p.prior.pred)
pred.prior.lower <- apply(p.prior.pred, 2, quantile, prob=0.025)
pred.prior.upper <- apply(p.prior.pred, 2, quantile, prob=0.975)

plot(temp.pred, p.post.pred[1,], type="n", xlab="Temperature (standardized)",
     ylab="Detection probability", ylim=c(0, 1.3), col=gray(0.8))
polygon(x=c(temp.pred, rev(temp.pred)),
        y=c(pred.post.lower, rev(pred.post.upper)),
        col=rgb(0,0,1,0.5), border=NA)                   # Post CI
polygon(x=c(temp.pred, rev(temp.pred)),
        y=c(pred.prior.lower, rev(pred.prior.upper)),
        col=rgb(0,1,0,0.5), border=NA)                   # Prior CI
for(i in seq(1, n.iter, by=100)) {  ## Thin by 100
    lines(temp.pred, p.prior.pred[i,], col=gray(0.8))  } # Prior preds
lines(temp.pred, pred.post.mean, col="blue", lwd=2)
lines(temp.pred, pred.prior.mean, col="darkgreen", lwd=2)
legend(-3, 1.3, c("Prior mean", "Prior samples", "Posterior mean"),
       col=c("darkgreen", "grey", "blue"), lwd=2)
```










