---
title: "HW_04"
author: "Henry Traynor"
date: 'Due: 5:00pm, 15 September'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a self-contained R script or Rmarkdown/Quarto file to do the following:

Using the simulated data, compare the prior and posterior predictive distributions of $\lambda$ for each of the 3 forest types. 

Hint: because forest type is categorical (not continuous), you can visualize the prior and posterior predictive distributions as histograms for each forest type.

```{r Data Generation}
set.seed(314159) # for reproducibility

nSites <- 100
nVisits <- 4

forest <- factor(sample(c("Hardwood", "Mixed", "Pine"), nSites, replace=TRUE))
forestMixed <- ifelse(forest=="Mixed", 1, 0)        ## Dummy
forestPine <- ifelse(forest=="Pine", 1, 0)          ## Dummy
temp <- matrix(rnorm(nSites*nVisits), nrow=nSites, ncol=nVisits)

# true coeffs
beta0 <- 0
beta1 <- -1
beta2 <- 1
lambda <- exp(beta0 + beta1*forestMixed + beta2*forestPine)
alpha0 <- -2
alpha1 <- 1
p <- plogis(alpha0 + alpha1*temp)

# true lambda plots
lambda.true <- c(
  Hardwood = exp(beta0),
  Mixed = exp(beta0 + beta1),
  Pine = exp(beta0 + beta2)
)
lambda.true

# simulate occupancy and detection data
N <- rpois(nSites, lambda=lambda)         ## local abundance 
y <- matrix(NA, nrow=nSites, ncol=nVisits)
for(i in 1:nSites) {
    y[i,] <- rbinom(nVisits, size=N[i], prob=p[i,])
}

# summary statistics
maxCounts <- apply(y, 1, max) 
table(maxCounts)            

naiveOccupancy <- sum(maxCounts>0)/nSites
naiveOccupancy 

naiveAbund <- sum(maxCounts)
naiveAbund
```

```{r Bayesian Modeling}
# prep data in named list
jags.data <- list(y=y,
                  forestMixed=forestMixed,
                  forestPine=forestPine,
                  nSites=nSites, 
                  temp=temp,
                  nOccasions=nVisits)
# initial values for MCMC
jags.inits <- function() {
    list(beta0=rnorm(1), alpha0=rnorm(1), N=maxCounts)
}
# parameters to save
jags.pars <- c("beta0", "beta1", "beta2",
               "alpha0", "alpha1", "totalAbundance")

# run model
library(jagsUI)
jags.post.samples <- jags.basic(data=jags.data, 
                                inits=jags.inits,
                                parameters.to.save=c(jags.pars,"N"), 
                                model.file="Nmix-model-covs.jag",
                                n.chains=3,
                                n.adapt=100, 
                                n.burnin=0,
                                n.iter=6000,
                                parallel=TRUE)
```

```{r Summary & Plots}
summary(jags.post.samples[,jags.pars])
plot(jags.post.samples[,paste0("N[", 1:2, "]")])
plot(jags.post.samples[,paste0("N[", 3:4, "]")])
plot(jags.post.samples[,jags.pars[1:3]])
plot(jags.post.samples[,jags.pars[4:5]])
```

```{r Bayesian Prediction}
# extract beta values
lambda.coef.post <- as.matrix(jags.post.samples[,c("beta0","beta1", "beta2")])
head(lambda.coef.post, n=4)
n.iter <- nrow(lambda.coef.post)
# vector for predictive val
forest.pred <- c(rep("Hardwood", 50),
                 rep("Mixed", 50),
                 rep("Pine",50)
                 )

forest.Mixed.pred <- ifelse(forest.pred=="Mixed", 1, 0)
forest.Pine.pred <- ifelse(forest.pred=="Pine", 1, 0)

lambda.post.pred <- matrix(NA, nrow=n.iter, ncol=length(forest.pred))

# we now use exp() rather than plogis() since lambda is log-linear rather than logit-linear
for(i in 1:n.iter) {
    lambda.post.pred[i,] <- exp(
      lambda.coef.post[i,"beta0"] + 
      lambda.coef.post[i,"beta1"] * forest.Mixed.pred + 
      lambda.coef.post[i,"beta2"] * forest.Pine.pred
      )
}

# plot showing forest type on categorical horiz. axis and lambda on continuous vert. vertical axis
# pull out values for each forest type
hardwood <- as.vector(lambda.post.pred[, 1:50])
mixed    <- as.vector(lambda.post.pred[, 51:100])
pine     <- as.vector(lambda.post.pred[, 101:150])
#combine vectors in long form for easy graphing
lambda.long <- data.frame(
  lambda = c(hardwood, mixed, pine),
  forest = rep(
    c("Hardwood", "Mixed", "Pine"),
    times = c(length(hardwood), length(mixed), length(pine))
  )
)
library(ggplot2)
ggplot(lambda.long, aes(x = lambda)) +
  geom_histogram() +
  facet_wrap(~forest, ncol = 3) +
  labs(
    x = "lambda_hat",
    y = "Frequency"
  ) + 
  geom_vline(
    data = data.frame(
      forest = names(lambda.true),
      lambda = lambda.true
    ),
    aes(xintercept = lambda, color = "True value"),
    linetype = "solid"
)
```

```{r Prior Sampling}
jags.data2 <- jags.data # redefine data
jags.data2$y[] <- NA     # Replace data with missing values

# resample
jags.prior.samples <- jags.basic(data=jags.data2,
                                 inits=jags.inits,
                                 parameters.to.save=jags.pars,
                                 model.file="Nmix-model-covs.jag",
                                 n.chains=3,
                                 n.adapt=100,
                                 n.burnin=0,
                                 n.iter=6000,
                                 parallel=TRUE)
summary(jags.prior.samples)
```

```{r Prior Prediction}
lambda.coef.prior <- as.matrix(jags.prior.samples[,c("beta0","beta1","beta2")])
lambda.prior.pred <- matrix(NA, nrow=n.iter, ncol=length(forest.pred))
for(i in 1:n.iter) {
    lambda.prior.pred[i,] <- exp(lambda.coef.prior[i,"beta0"] + lambda.coef.prior[i,"beta1"]*forest.Mixed.pred + lambda.coef.prior[i, "beta2"]*forest.Pine.pred)
}

# plot showing forest type on categorical horiz. axis and lambda on continuous vert. vertical axis
hardwood.prior <- as.vector(lambda.prior.pred[, 1:50])
mixed.prior    <- as.vector(lambda.prior.pred[, 51:100])
pine.prior     <- as.vector(lambda.prior.pred[, 101:150])
lambda.long.prior <- data.frame(
  lambda = c(hardwood.prior, mixed.prior, pine.prior),
  forest = rep(
    c("Hardwood", "Mixed", "Pine"),
    times = c(
      length(hardwood.prior),
      length(mixed.prior),
      length(pine.prior)
    )
  )
)

log.lambda.long.prior <- lambda.long.prior
log.lambda.long.prior$lambda <- log(log.lambda.long.prior$lambda)
ggplot(log.lambda.long.prior, aes(x = lambda)) +
  geom_histogram() +
  facet_wrap(~forest, ncol = 3) +
  labs(
    x = "log(lambda_hat)",
    y = "Frequency"
  ) + 
  geom_vline(
    data = data.frame(
      forest = names(lambda.true),
      lambda = lambda.true
    ),
    aes(xintercept = lambda, color = "True value"),
    linetype = "solid"
  )
```

```{r Comparison Plot}
lambda.long$distribution <- "Posterior"
log.lambda.long.prior$distribution <- "Prior"
lambda.combined <- rbind(lambda.long, log.lambda.long.prior)

ggplot(lambda.combined,
       aes(x = lambda, fill = distribution)) +
  geom_histogram(
    position = "identity",
    alpha = 0.5,
  ) +
  facet_wrap(~ forest, ncol = 3) +
  labs(
    x = "Posterior: lambda_hat \n Prior: log(lambda_hat)",
    y = "Frequency",
    fill = "Distribution"
  ) +
  geom_vline(
      data = data.frame(
        forest = names(lambda.true),
        lambda = lambda.true
      ),
      aes(xintercept = lambda, color = "True value, original scale"),
      linetype = "solid"
  )
```
Note:

I have showed the posterior $\lambda$ distribution on it's original scale. Due to the relatively much wider prior distribution, I have displayed the prior $\ambda$ distribution on the log-scale, as shown in the horizontal axis label. The vertical line showing the true value of $\lambda$ is on the original scale and thus is only applicable when visually expecting the accuracy of the posterior distribution, not the prior distribution.










