---
title: "HW_06"
output:
  pdf_document: default
  html_document: default
date: 'Due: 5:00pm, 29 September'
---

## Assignment

Create a self-contained R script or Rmarkdown file to do the following:

- Fit a distance sampling model with a half-normal detection function and the following covariates to the black-throated blue warbler data (`btbw_data_distsamp.csv`) in `unmarked` and `JAGS`:
  - Density covariates: `Elevation`, `UTM.N`, `UTM.W`
  - Detection covariates: `Wind`, `Noise`
  - Response: `btbw0_20`, `btbw20_40`, `btbw40_60`, `btbw60_80`, `btbw80_100`
- Using the model fitted in `unmarked`, create two graphs of the predictions: one for density and the other for the scale parameter ($\sigma$).

### Setup
```{r setup}
library(dplyr)
data <- read.csv("btbw_data_distsamp.csv")
data <- data %>%
  rename(
    btbw0_20 = X.0.20.,
    btbw20_40 = X.20.40.,
    btbw40_60 = X.40.60.,
    btbw60_80 = X.60.80.,
    btbw80_100 = X.80.100.,
  )

head(data)
```

### `unmarked`
```{r unmarked}
library(unmarked)

# extract covariates and standardize
covs <- data.frame(Elevation = data$Elevation, 
                       UTM_N = data$UTM.N, 
                       UTM_W = data$UTM.W,
                       Wind = data$Wind, 
                       Noise = data$Noise
                       )

covs_standard <- covs %>%
  mutate(across(where(is.numeric), ~ as.vector(scale(.))))

# extract response var as matrix
y <- data %>%
  select(btbw0_20,
         btbw20_40,
         btbw40_60,
         btbw60_80,
         btbw80_100)
y <- as.matrix(y)

# 'dist.breaks' for 'unmarkedFrameDS'
b <- c(0, 20, 40, 60, 80, 100)


# fit model in 'unmarked'
umf <- unmarkedFrameDS(y = y,
                       siteCovs = covs_standard,
                       dist.breaks = b,
                       survey = "point",
                       unitsIn = "m")

# display summary
summary(umf)

fm <- distsamp(~Noise+Wind ~ UTM_N+UTM_W+Elevation, umf, keyfun = "halfnorm")
fm
```

### `JAGS`

#### State model (Poisson)
$$\textrm{log}(\lambda_i) = \beta_0+\beta_1 w_{i1}+\beta_2 w_{i2}+\dots$$
$$N_i \sim \textrm{Poisson}(\lambda_i)$$

#### Observation Model
$$\textrm{log}(\sigma_i) = \alpha_0 + \alpha_1 w_{i1} + \alpha_2 w_{i3} + \dots$$
$$\{ y_{i1}, \dots, y_{iK} \} \sim \textrm{Multinomial}(N_i, \pi(x, \sigma_i))$$
```{r JAGS}
model_string <- "
model {

beta0 ~ dnorm(0, 0.1)
beta1 ~ dnorm(0, 0.1)
beta2 ~ dnorm(0, 0.1)
beta3 ~ dnorm(0, 0.1)

alpha0 ~ dnorm(0, 0.1)
alpha1 ~ dnorm(0, 0.1)
alpha2 ~ dnorm(0, 0.1)

for(i in 1:nSites) {
  log(lambda[i]) <- beta0 + beta1*Elevation[i] + beta2*UTM_N[i] + beta3*UTM_W[i]
  N[i] ~ dpois(lambda[i]*Area)
  
  
  log(sigma[i]) <- alpha0 + alpha1*Noise[i] + alpha2*Wind[i]
  for(j in 1:nBins) {
    pbar[i,j] <- (sigma[i]^2 * (1-exp(-b[j+1]^2/(2*sigma[i]^2))) -
                  sigma[i]^2 * (1-exp(-b[j]^2/(2*sigma[i]^2)))) * 
                  2*3.141593/area[j]
    pi[i,j] <- psi[j]*pbar[i,j]
  }
  pi[i,nBins+1] <- 1-sum(pi[i,1:nBins]) ## Pr(not detected)
  n[i] ~ dbin(1-pi[i,nBins+1], N[i])
  y[i,] ~ dmulti(pi[i,1:nBins]/(1-pi[i,nBins+1]), n[i])
  ## If N~Pois(lam), then the above is equivalent to:
  # for(j in 1:nBins) { y[i,j] ~ dpois(lambda[i]*pi[i,j])  }
}

totalAbundance <- sum(N[1:nSites])

}
"
# make tempfile so model string can be referenced without file
model_file <- tempfile(fileext = ".jags")
writeLines(model_string,
           con = model_file
           )

B <- 100
Area <- base::pi *B^2/1e4
nSites <- nrow(y)
J <- ncol(y)
b <- seq(0,B, length=J+1)
area <- base::pi *b^2
psi <- (area[-1]-area[-(J+1)]) / area[J+1]

jags_data <- list(y=y, 
                  n=rowSums(y), 
                  area=diff(area),
                  b=b,              # Distance break points
                  Area=Area,  # Area in ha
                  psi=psi,          # Pr(occuring in bin j)
                  Elevation=covs_standard$Elevation, 
                  UTM_N=covs_standard$UTM_N,
                  UTM_W=covs_standard$UTM_W,
                  Noise=covs_standard$Noise,
                  Wind=covs_standard$Wind, 
                  nSites=nSites, 
                  nBins=J
                  )

jags_inits <- function() {
  list(beta0 = rnorm(1),
       alpha0 = log(20),
       N = rowSums(y) + rpois(nrow(y),2)
       )
}

jags_pars <- c("beta0",
                "beta1",
                "beta2",
                "beta3",
                "alpha0",
                "alpha1",
                "alpha2",
                "totalAbundance"
                )



library(jagsUI)
jags_post <- jags.basic(data=jags_data,
                        inits = jags_inits,
                        parameters.to.save = jags_pars,
                        model.file = model_file,
                        n.chains = 3,
                        n.adapt = 100,
                        n.burnin = 1000,
                        n.iter = 5000,
                        parallel = TRUE)

round(summary(jags_post)$quantile, digits=3)

plot(jags_post[,jags_pars[1]], main = jags_pars[1])
plot(jags_post[,jags_pars[2]], main = jags_pars[2])
plot(jags_post[,jags_pars[3]], main = jags_pars[3])
plot(jags_post[,jags_pars[4]], main = jags_pars[4])
plot(jags_post[,jags_pars[5]], main = jags_pars[5])
plot(jags_post[,jags_pars[6]], main = jags_pars[6])
plot(jags_post[,jags_pars[7]], main = jags_pars[7])
plot(jags_post[,jags_pars[8]], main = jags_pars[8])
```
### `unmarked` prediction

```{r}
library(ggplot2)

#density
pred_data <- data.frame(Elevation = seq(min(fm@data@siteCovs$Elevation),
                                        max(fm@data@siteCovs$Elevation),
                                        length.out = 100
                                        ),
                        UTM_N = mean(fm@data@siteCovs$UTM_N),
                        UTM_W = mean(fm@data@siteCovs$UTM_W)
                        )

pred_density <- predict(
  fm,
  type = "state",
  newdata = pred_data
)

pred_data$Density <- pred_density$Predicted
pred_data$lower <- pred_density$lower
pred_data$upper <- pred_density$upper

ggplot(pred_data, aes(x=Elevation, y=Density)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5) +
  geom_line(linewidth=1) +
  labs(
    x = "Elevation",
    y = "Predicted Density at mean coordinates"
  ) +
  ggtitle("Predicted Density")

pred_sigma_data <- data.frame(
  Noise = seq(min(fm@data@siteCovs$Noise),
              max(fm@data@siteCovs$Noise),
              length.out = 100
              ),
  Wind = mean(fm@data@siteCovs$Wind)
)

pred_sigma <- predict(
  fm,
  type="det",
  newdata = pred_sigma_data
)

pred_sigma_data$sigma <- pred_sigma$Predicted
pred_sigma_data$lower <- pred_sigma$lower
pred_sigma_data$upper <- pred_sigma$upper

ggplot(pred_sigma_data, aes(x=Noise, y=sigma)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5) +
  geom_line(linewidth=1) +
  labs(
    x = "Noise",
    y = "Predicted sigma at mean wind"
  ) +
  ggtitle("Predicted sigma")
```