---
title: "HW_03"
author: "Henry Traynor"
date: 'Due: 5:00pm, 8 September'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a self-contained R script or Rmarkdown/Quarto file to do the following:

1. Fit an occupancy model in ‘unmarked’ to the Ruffed Grouse data (`"grouse_data.csv"`). Use all three covariates, with elevation and forest cover affecting occupancy and day of the year affecting detection probability.

* Response: `grouse1`, `grouse2`, `grouse3`
* Site covs: Elevation (`elev`), forest cover (`fc`)
* Obs cov: Day of the year (`doy.1`, `doy.2`, `doy.3`)

```{r Data import}
setwd("C:/Users/henry/Desktop/WILD8390/Homework/HW03")
grouse_data <- read.csv("grouse_data.csv")
```

```{r unmarked}
library(unmarked)
umf <-unmarkedFrameOccu(y = grouse_data[,c("grouse1", "grouse2", "grouse3")],
                        siteCovs = grouse_data[,c("elev", "fc")],
                        obsCovs = list(doy=grouse_data[,c("doy.1","doy.2","doy.3")])
                        )
summary(umf)

# standarize cont. variables
mean.elev <- mean(grouse_data[,"elev"])
mean.forest <- mean(grouse_data[,"fc"])
sd.elev <- sd(grouse_data[,"elev"])
sd.forest <- sd(grouse_data[,"fc"])
elev_s <- (grouse_data[,"elev"]-mean.elev)/sd.elev
forest_s <- (grouse_data[,"fc"]-mean.forest)/sd.forest
siteCovs(umf)$elev_s <- elev_s
siteCovs(umf)$forest_s <- forest_s

# fit the model
fm <- occu(~elev_s+doy ~elev_s+forest_s, umf) # Standardized covariates
fm

# prediction
pred.data <- data.frame(elev_s=seq(from=-3, to=3, length=100),
                        forest_s=0, doy=100) # Standardized
pred.data$elev <- pred.data$elev_s*sd.elev+mean.elev # Orig scale
pred.data$fc <- pred.data$forest_s*sd.forest+mean.forest 

psi.pred <- predict(fm, newdata=pred.data,
                    type='state', append=TRUE)
p.pred <- predict(fm, newdata=pred.data,
                  type='det', append=TRUE)

print(head(psi.pred, n=4), digits=2)
print(head(p.pred, n=4), digits=2)
```

2. Make one prediction graph for occupancy and one for detection.

```{r Plots: unmarked}
# these plots might not generate here but will if you knit it. PDF is also available at github link at the top of document
par(mai=c(0.9,0.9,0.1,0.1))  
plot(Predicted ~ elev_s, psi.pred, type="l", ylab="Occurrence probability",
     col="blue", xlab="Elevation (standardized)", ylim=0:1) 
lines(lower ~ elev_s, psi.pred, col="grey"); lines(upper ~ elev_s, psi.pred, col="grey")

par(mai=c(0.9,0.9,0.1,0.1))  
plot(Predicted ~ elev, psi.pred, type="l", ylab="Occurrence probability",
     col="blue", xlab="Elevation (original scale)", ylim=0:1)
lines(lower ~ elev, psi.pred, col="grey"); lines(upper ~ elev, psi.pred, col="grey")
```

3. Create the same graphs as above, but using Bayesian methods.

```{r Bayesian model}
nSites <- nrow(grouse_data)
nVisits <- 3

jags.data <- list(y=grouse_data[,c("grouse1", "grouse2", "grouse3")],
                  elev=elev_s,
                  forest=forest_s,
                  doy=grouse_data[,c("doy.1","doy.2","doy.3")],
                  nSites=nSites,
                  nOccasions=nVisits)

jags.inits <- function() {
    list(beta0=rnorm(1), alpha0=rnorm(1), z=rep(1, nSites))
}

jags.pars <- c("beta0", "beta1", "beta2",
               "alpha0", "alpha1", "alpha2", "alpha3",
               "sitesOccupied")

library(jagsUI)
library(coda)
jags.post.samples <- jags.basic(data=jags.data, inits=jags.inits,
                                parameters.to.save=jags.pars,
                                model.file="occupancy-model-covs.jag",
                                n.chains=3, n.adapt=100, n.burnin=0,
                                n.iter=2000, parallel=TRUE)

### ------------------------------------------
### RUN THESE LINES IN YOUR CONSOLE
### ------------------------------------------
png("jags_plot.png", width = 500, height = 800)
plot(jags.post.samples)

psi.coef.post <- as.matrix(jags.post.samples[,c("beta0","beta1","beta2")])
head(psi.coef.post, n=4)

n.iter <- nrow(psi.coef.post)  
psi.post.pred <- matrix(NA, nrow=n.iter, ncol=nrow(pred.data))

for(i in 1:n.iter) {
    psi.post.pred[i,] <- plogis(psi.coef.post[i,"beta0"] +
                                psi.coef.post[i,"beta1"]*pred.data$elev_s)
}

plot(pred.data$elev, psi.post.pred[1,], type="l", xlab="Elevation",
     ylab="Occurrence probability", ylim=c(0, 1), col=gray(0.8))

for(i in 1:n.iter) {
    lines(pred.data$elev, psi.post.pred[i,], col=gray(0.8))
}

pred.post.mean <- colMeans(psi.post.pred)
pred.post.lower <- apply(psi.post.pred, 2, quantile, prob=0.025)
pred.post.upper <- apply(psi.post.pred, 2, quantile, prob=0.975)
lines(pred.data$elev, pred.post.mean, col="blue")
lines(pred.data$elev, pred.post.lower, col="blue", lty=2)
lines(pred.data$elev, pred.post.upper, col="blue", lty=2)
```
