---
title: "HW_04"
author: "Henry Traynor"
date: 'Due: 5:00pm, 29 September'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a self-contained R script or Rmarkdown file to do the following:

Simulate independent double observer data with the following properties:
\begin{itemize}
  \item nSites=200
  \item $\lambda$=3
  \item $p_1$=0.3 and $p_2$=0.5
\end{itemize}

```{r Data Simulation}
set.seed(314159)

nSites <- 200
lambda <- 3

N <- rpois(nSites, lambda)

summary(N)
mean(N)
var(N)
```

```{r Obs Data}
p1 <- 0.3
p2 <- 0.5

pi1 <- p1*(1-p2)
pi2 <- (1-p1)*p2
pi3 <- p1*p2
pi4 <- (1-p1)*(1-p2)

cp <- c(pi1, 
        pi2, 
        pi3
        )

y <- matrix(NA,
            nrow=nSites,
            ncol=3
)

colnames(y) <- c("obs1_only", 
                 "obs2_only", 
                 "both"
                 )

for(i in 1:nSites) {
  counts <- rmultinom(1, size=N[i], prob = c(cp,pi4))
  y[i,] <- counts[1:3]
}

head(y)
summary(y)
```

Fit the model using `unmarked`. You will need to create an observation covariate indicating observer A and B. When you create the `unmarkedFrameMPois`, use `type=double`.

```{r unmarked}
observer <- matrix(
  c("A", "B"),
  nrow = nSites,
  ncol = 2,
  byrow = TRUE
)
colnames(observer) <- c(
  "observer1",
  "observer2"
)

library(unmarked)
umf <- unmarkedFrameMPois(
  y = y,
  obsCovs = list(observer = observer),
  type = "double"
)

fm <- multinomPois(
  ~ observer -1 ~ 1,
  data = umf
)

fm

coef_estimates <- data.frame(
  parameter = c("lambda_hat", "pA_hat", "pB_hat"),
  true_value = c(lambda, p1, p2)
)

coef_estimates$unmarked <- c(exp(coef(fm)["lambda(Int)"]),
                             plogis(coef(fm)["p(observerA)"]),
                             plogis(coef(fm)["p(observerB)"])
                             )

coef_estimates
```

Fit the same model in JAGS

```{r JAGS}
library(jagsUI)

model_string <- "
model {

  lambda ~ dunif(0,20)
  pA ~ dunif(0,1)
  pB ~ dunif(0,1)
  
  pi1 <- pA * (1 - pB)
  pi2 <- (1 - pA) * pB
  pi3 <- pA * pB

  for (i in 1:nSites) {
    y[i,1] ~ dpois(lambda * pi1)
    y[i,2] ~ dpois(lambda * pi2)
    y[i,3] ~ dpois(lambda * pi2)
  }
}
"

model_file <- tempfile(fileext = ".jags")

writeLines(
  model_string,
  con = model_file
)

jags_data <- list(
  nSites = nSites,
  y = y
)

jags_inits <- function() {
  list(
    lambda = runif(1),
    pA = runif(1),
    pB = runif(1),
    N=rowSums(y)+rpois(nrow(y), 2)
  )
}

jags_pars <- c("lambda", "pA", "pB")

jags_post <- jags.basic(
  data = jags_data,
  inits = jags_inits,
  parameters.to.save = jags_pars,
  model.file = model_file,
  n.chains = 3,
  n.adapt = 100,
  n.burnin = 2000,
  n.iter = 10000,
  parallel = TRUE
)

jags_summary <- summary(jags_post)

coef_estimates$JAGS_mean <- jags_summary$statistics[c("lambda", "pA", "pB"), "Mean"]
coef_estimates$JAGS_lower95 <- jags_summary$quantiles[c("lambda", "pA", "pB"), "2.5%"]
coef_estimates$JAGS_median <- jags_summary$quantiles[c("lambda", "pA", "pB"), "50%"]
coef_estimates$JAGS_upper95 <- jags_summary$quantiles[c("lambda", "pA", "pB"), "97.5%"]
```
    
For both unmarked and JAGS, use the point estimates of $p_1$ and $p_2$ to compute the $\pi$ probabilities.

```{r pi values}
pi_estimates <- data.frame(
  parameter = c("pi1", "pi2", "pi3", "pi4"),
  
)
```
