---
title: "HW_07"
output:
  html_document: default
  pdf_document: default
date: 'Due: 5:00pm, 6 October'
---

# Assignment

Create a self-contained R script or Rmarkdown file to do the following:

Import the Canada warbler mark-recapture data `CAWAcaps_2018.csv` and fit the following models in `RMark` and `JAGS`:

* Model $M_0$
* Model $M_t$
* Model $M_h$ – Finite mixture
* Model $M_h$ – Logit-normal (`JAGS` only)

For `JAGS`, use $M=800$, but note if the posterior of $N$ pushes up against $M$.

Create a table reporting point estimates and 95% CIs for $N$ for each model fit in `RMark` and `JAGS`. Which model has the lowest AICc (`RMark` only)?

### Setup

```{r}
library(RMark)
library(jagsUI)

y <- read.csv("CAWAcaps_2018.csv")

nIndividuals <- nrow(y)
nOccasions <- ncol(y)

y.ch <- data.frame(ch=apply(y, 1, paste, collapse=""))

head(y.ch)
```

## `RMark` Model Fitting

```{r Model M0}
# capture Pr = recapture Pr
# thus 'share=TRUE'
mark_M0 <- mark(
  data = y.ch,
  model = "Closed",
  silent= TRUE,
  model.parameters = list(p = list(formula=~1, share=TRUE))
)
```

```{r Model Mt}
# unique p for each capture occasion
# 'formula =~time'
mark_Mt <- mark(
  data = y.ch,
  model = "Closed",
  silent = TRUE,
  model.parameters = list(p = list(formula=~time, share=TRUE))
)
```

```{r Model Mh: finite mixture}
# individual heterogeneity in p
# assumes a number of groups with different p
# 'formula=~mixture'
mark_Mh_mixture <- mark(
  data = y.ch,
  silent = TRUE,
  model = "HetClosed",
  model.parameters = list(p=list(formula=~mixture, share=TRUE))
)
```

## `JAGS` Model Fitting

Use $M=800$ to add $M-\texttt{nrow(y)}$ extra `0000` captures.

```{r Data Augmentation}
M <- 10000

y_aug <- matrix(0, M, nOccasions)
y_aug[1:nrow(y),] <- as.matrix(y)
```

```{r Model M0}
model_string_M0 <- "
model {
 p ~ dunif(0,1)
 psi ~ dunif(0,1)
 
 for(i in 1:M) {
  z[i] ~ dbern(psi)
  for(j in i:J) {
    y[i,j] ~ dbern(z[i]*p)
  }
 }
 N <- sum(z)
}
"

model_file_M0 <- tempfile(fileext = ".jags")
writeLines(model_string_M0,
           con = model_file_M0
           )

jags_data_M0 <- list(y=y_aug, M=M, J=nOccasions)
jags_inits_M0 <- function () list(z=rep(1,M), psi=runif(1), p=runif(1))
jags_pars_M0 <- c("p", "psi", "N")

jags_post_M0 <- jags.basic(
  data = jags_data_M0,
  inits = jags_inits_M0,
  parameters.to.save = jags_pars_M0,
  model.file = model_file_M0,
  n.chains = 3,
  n.adapt = 1000,
  n.burnin = 0,
  n.iter = 20000,
  parallel = TRUE
)
```

## Point Estimates and AICc Values

plot(jags_post_M0)

