---
title: "HW_07"
output:
  html_document: default
  pdf_document: default
date: 'Due: 5:00pm, 6 October'
---

# Assignment

Create a self-contained R script or Rmarkdown file to do the following:

Import the Canada warbler mark-recapture data `CAWAcaps_2018.csv` and fit the following models in `RMark` and `JAGS`:

* Model $M_0$
* Model $M_t$
* Model $M_h$ – Finite mixture
* Model $M_h$ – Logit-normal (`JAGS` only)

For `JAGS`, use $M=800$, but note if the posterior of $N$ pushes up against $M$.

Create a table reporting point estimates and 95% CIs for $N$ for each model fit in `RMark` and `JAGS`. Which model has the lowest AICc (`RMark` only)?

### Setup

```{r}
library(RMark)
library(jagsUI)

y <- read.csv("CAWAcaps_2018.csv")

nIndividuals <- nrow(y)
nOccasions <- ncol(y)

y.ch <- data.frame(ch=apply(y, 1, paste, collapse=""))

head(y.ch)
```

## `RMark` Model Fitting

```{r Model M0}
# capture Pr = recapture Pr
# thus 'share=TRUE'
mark_M0 <- mark(
  data = y.ch,
  model = "Closed",
  silent= TRUE,
  model.parameters = list(p = list(formula=~1, share=TRUE))
)
```

```{r Model Mt}
mark_Mt <- mark(
  data = y.ch,
  model = "Closed",
  silent = TRUE,
  model.parameters = list(p = list(formula=~time, share=TRUE))
)
```

```{r Model Mh: finite mixture}
mark_Mh_mixture <- mark(
  data = y.ch,
  silent = TRUE,
  model = "HetClosed",
  model.parameters = list(p=list(formula=~mixture, share=TRUE))
)
```

## `JAGS` Model Fitting

## Point Estimates and AICc Values