---
title: "HW_07"
output:
  pdf_document: default
  html_document: default
date: 'Due: 5:00pm, 6 October'
---

# Assignment

Create a self-contained R script or Rmarkdown file to do the following:

Import the Canada warbler mark-recapture data `CAWAcaps_2018.csv` and fit the following models in `RMark` and `JAGS`:

* Model $M_0$
* Model $M_t$
* Model $M_h$ – Finite mixture
* Model $M_h$ – Logit-normal (`JAGS` only)

For `JAGS`, use $M=800$, but note if the posterior of $N$ pushes up against $M$.

Create a table reporting point estimates and 95% CIs for $N$ for each model fit in `RMark` and `JAGS`. Which model has the lowest AICc (`RMark` only)?

### Setup

```{r}
library(RMark)
library(jagsUI)

y <- read.csv("CAWAcaps_2018.csv")

nIndividuals <- nrow(y)
nOccasions <- ncol(y)
J <- nOccasions

y.ch <- data.frame(ch=apply(y, 1, paste, collapse=""))

head(y.ch)
```

## `RMark` Model Fitting

```{r Model M0-Mark}
# capture Pr = recapture Pr
# thus 'share=TRUE'
mark_M0 <- mark(
  data = y.ch,
  model = "Closed",
  silent= FALSE,
  model.parameters = list(p = list(formula=~1, share=TRUE))
)
```

```{r Model Mt-Mark}
# unique p for each capture occasion
# 'formula =~time'
mark_Mt <- mark(
  data = y.ch,
  model = "Closed",
  silent = TRUE,
  model.parameters = list(p = list(formula=~time, share=TRUE))
)
```

```{r Model Mh- finite mixture:Mark}
# individual heterogeneity in p
# assumes a number of groups with different p
# 'formula=~mixture'
mark_Mh_mixture <- mark(
  data = y.ch,
  silent = TRUE,
  model = "HetClosed",
  model.parameters = list(p=list(formula=~mixture, share=TRUE))
)
```

## `JAGS` Model Fitting

Use $M=800$ to add $M-\texttt{nrow(y)}$ extra `0000` captures.

```{r Data Augmentation}
M <- 800

y_aug <- matrix(0, M, nOccasions)
y_aug[1:nrow(y),] <- as.matrix(y)
```

```{r Model M0-JAGS}
model_string_M0 <- "
model {

p ~ dunif(0,1)             ## Capture probability
psi ~ dunif(0,1)           ## Data augmentation parameter

for(i in 1:M) {
  z[i] ~ dbern(psi)        ## DA indicator
  for(j in 1:J) {
    y[i,j] ~ dbern(z[i]*p) ## Model for the capture histories
  }
}

N <- sum(z)                ## Abundance

}
"

model_file_M0 <- tempfile(fileext = ".jags")
writeLines(model_string_M0,
           con = model_file_M0
           )

jags_data_M0 <- list(y=y_aug, M=M, J=nOccasions)
jags_inits_M0 <- function () list(z=rep(1,M), psi=runif(1), p=runif(1))
jags_pars_M0 <- c("p", "psi", "N")

jags_post_M0 <- jags.basic(
  data = jags_data_M0,
  inits = jags_inits_M0,
  parameters.to.save = jags_pars_M0,
  model.file = model_file_M0,
  n.chains = 3,
  n.adapt = 1000,
  n.burnin = 0,
  n.iter = 2000,
  parallel = TRUE
)

plot(jags_post_M0)
```

```{r Model Mt - JAGS}
model_string_Mt <- "
model {
p1 ~ dunif(0,1)
p2 ~ dunif(0,1)
p3 ~ dunif(0,1)
p4 ~ dunif(0,1)
psi ~ dunif(0,1)

p <- c(p1, p2, p3, p4)

for(i in 1:M) {
  z[i] ~ dbern(psi)
  for(j in 1:J) {
    y[i,j] ~ dbern(z[i]*p[j])
  }
}
N <- sum(z)
}
"

model_file_Mt <- tempfile(fileext = ".jags")
writeLines(model_string_Mt,
           con = model_file_Mt
           )

jags_data_Mt <- list(y=y_aug, M=M, J=nOccasions)
jags_inits_Mt <- function () list(z=rep(1,M),
                                  psi=runif(1),
                                  p1=runif(1),
                                  p2=runif(1),
                                  p3=runif(1),
                                  p4=runif(1)
                                  )
jags_pars_Mt <- c("p1","p2","p3","p4", "psi", "N")

jags_post_Mt <- jags.basic(
  data = jags_data_Mt,
  inits = jags_inits_Mt,
  parameters.to.save = jags_pars_Mt,
  model.file = model_file_Mt,
  n.chains = 3,
  n.adapt = 1000,
  n.burnin = 0,
  n.iter = 2000,
  parallel = TRUE
)

# call plots like this to avoid figure margin errors
plot(jags_post_Mt[,jags_pars_Mt[1]], main = jags_pars_Mt[1])
plot(jags_post_Mt[,jags_pars_Mt[2]], main = jags_pars_Mt[2])
plot(jags_post_Mt[,jags_pars_Mt[3]], main = jags_pars_Mt[3])
plot(jags_post_Mt[,jags_pars_Mt[4]], main = jags_pars_Mt[4])
plot(jags_post_Mt[,jags_pars_Mt[5]], main = jags_pars_Mt[5])
plot(jags_post_Mt[,jags_pars_Mt[6]], main = jags_pars_Mt[6])

```


Did this one not realizing it wasn't required. But leaving it here for my own notes.

```{r Model Mb-JAGS}
model_string_Mb <- "
model {
p ~ dunif(0, 1)     ## Capture prob
c ~ dunif(0, 1)     ## Recapture prob

psi ~ dunif(0,1)    ## Data augmentation parameter

for(i in 1:M) {     
  z[i] ~ dbern(psi) ## DA indicator
  for(j in 1:J) {
    pc[i,j] <- ifelse(prevcap[i,j]==0, p, c) ## First cap or recap?
    y[i,j] ~ dbern(z[i]*pc[i,j])
  }
}

N <- sum(z)
}
"
model_file_Mb <- tempfile(fileext = ".jags")
writeLines(model_string_Mb,
           con = model_file_Mb
           )

jags_data_Mb <- jags_data_M0

prevcap_aug <- matrix(FALSE, M, J)
for(i in 1:M) {
  for(j in 2:J) {
    prevcap_aug[i,j] <- any(y_aug[i,1:(j-1)]>0)
  }
}

jags_data_Mb$prevcap <- ifelse(prevcap_aug, 1, 0)
jags_inits_Mb <- function () {
  list(z=rep(1,M),psi=runif(1), p=runif(1), c=runif(1)) 
}


jags_pars_Mb <- c("p", "c", "psi", "N")

jags_post_Mb <- jags.basic(
  data = jags_data_Mb,
  inits = jags_inits_Mb,
  parameters.to.save = jags_pars_Mb,
  model.file = model_file_Mb,
  n.chains = 3,
  n.adapt = 1000,
  n.burnin = 0,
  n.iter = 2000,
  parallel = TRUE
)

plot(jags_post_Mb)
```

```{r Model Mh- finite mixture-JAGS}
model_string_Mh_finite <- "
model {

p1 ~ dunif(0, 1)     ## Capture prob, group 1
p2 ~ dunif(0, 1)     ## Capture prob, group 2
mixprob ~ dunif(0,1) ## Finite mixture prob
psi ~ dunif(0,1)     ## Data augmentation parameter

for(i in 1:M) {     
  z[i] ~ dbern(psi)  ## DA indicator
  group[i] ~ dbern(mixprob)           ## Group 1 or 2?
  p[i] <- ifelse(group[i]==0, p1, p2) ## p depends on group
  for(j in 1:J) {
    y[i,j] ~ dbern(z[i]*p[i])
  }
}
N <- sum(z)
}
"

model_file_Mh_finite <- tempfile(fileext = ".jags")
writeLines(model_string_Mh_finite,
           con = model_file_Mh_finite
           )

jags_data_Mh_finite <- list(y=y_aug, M=M, J=nOccasions)
jags_inits_Mh_finite <- function () list(z=rep(1,M),
                                         psi=runif(1),
                                         p1=runif(1),
                                         p2=runif(1),
                                         mixprob=runif(1)
                                         )
jags_pars_Mh_finite <- c("p1","p2","mixprob", "psi", "N")

jags_post_Mh_finite <- jags.basic(
  data = jags_data_Mh_finite,
  inits = jags_inits_Mh_finite,
  parameters.to.save = jags_pars_Mh_finite,
  model.file = model_file_Mh_finite,
  n.chains = 3,
  n.adapt = 1000,
  n.burnin = 0,
  n.iter = 2000,
  parallel = TRUE
)

plot(jags_post_Mh_finite)
```

```{r Model Mh - logit normal}
model_string_Mh_logit <- "
model {

lp.mean ~ dnorm(0,0.1)   ## Mean of logit(p)
lp.var ~ dexp(1)         ## Variance of logit(p)
psi ~ dunif(0,1)         ## Data augmentation parameter

for(i in 1:M) {     
  z[i] ~ dbern(psi)      ## DA indicator
  lp[i] ~ dnorm(lp.mean, 1/lp.var) ## Random effect, p on logit
  p[i] <- ilogit(lp[i])  ## Capture prob for individual i
  for(j in 1:J) {
    y[i,j] ~ dbern(z[i]*p[i])
  }
}

N <- sum(z)

}
"

model_file_Mh_logit <- tempfile(fileext = ".jags")
writeLines(model_string_Mh_logit,
           con = model_file_Mh_logit
           )

jags_data_Mh_logit <- list(y=y_aug, M=M, J=nOccasions)
jags_inits_Mh_logit <- function () list(z=rep(1,M),
                                        psi=runif(1),
                                        lp.mean=runif(1),
                                        lp.var=runif(1)
                                        )
jags_pars_Mh_logit <- c("lp.mean","lp.var", "psi", "N")

jags_post_Mh_logit <- jags.basic(
  data = jags_data_Mh_logit,
  inits = jags_inits_Mh_logit,
  parameters.to.save = jags_pars_Mh_logit,
  model.file = model_file_Mh_logit,
  n.chains = 3,
  n.adapt = 1000,
  n.burnin = 0,
  n.iter = 2000,
  parallel = TRUE
)

plot(jags_post_Mh_logit)
```

## Point Estimates and AICc Values

```{r Mark Estimates}
mark_fits <- list(mark_M0,
                  mark_Mt,
                  mark_Mh_mixture
                  )

mark_estimates <- data.frame(
  Model = c("M0",
            "Mt",
            "Mh_finite"),
  N_Estimate = vector("numeric", length = 3),
  N_Upper = vector("numeric", length = 3),
  N_Lower = vector("numeric", length = 3),
  AICc = vector("numeric", length = 3)
)

for(i in 1:length(mark_fits)) {
  mark_estimates$N_Estimate[i] <- mark_fits[[i]]$results$derived$'N Population Size'$estimate
  mark_estimates$N_Upper[i] <- mark_fits[[i]]$results$derived$'N Population Size'$ucl
  mark_estimates$N_Lower[i] <- mark_fits[[i]]$results$derived$'N Population Size'$lcl
  mark_estimates$AICc[i] <- mark_fits[[i]]$results$AICc
}
```

```{r JAGS Estimates}
JAGS_posts <- list(
  jags_post_M0,
  jags_post_Mt,
  jags_post_Mh_finite,
  jags_post_Mh_logit
)


JAGS_estimates <- data.frame(
  Model = c("M0", "Mt", "Mh_finite", "Mh_logit"),
  N_Estimate = vector("numeric", length = 4),
  N_Upper = vector("numeric", length = 4),
  N_Lower = vector("numeric", length = 4)
)

for(i in 1:length(JAGS_posts)) {
  j_summary <- summary(JAGS_posts[[i]])
  JAGS_estimates$N_Estimate[i] <- j_summary$statistics["N","Mean"]
  JAGS_estimates$N_Upper[i] <- j_summary$quantiles["N","97.5%"]
  JAGS_estimates$N_Lower[i] <- j_summary$quantiles["N","2.5%"]
}
```

```{r Figures}
library(ggplot2)
library(gridExtra)
library(ggbreak)
mark_plot <- ggplot(data = mark_estimates, aes(x=Model, y=N_Estimate)) +
  geom_col() +
  geom_errorbar(
    aes(ymin = N_Lower, ymax = N_Upper),
    width = 0.2
  ) +
  labs(
    x = "Model",
    y = "Estimated population size"
  ) +
  scale_y_break(c(750,4000), ticklabels = c(4000, 4500))

JAGS_plot <- ggplot(data = JAGS_estimates, aes(x=Model, y=N_Estimate)) +
  geom_col() +
  geom_errorbar(
    aes(ymin = N_Lower, ymax = N_Upper),
    width = 0.2
  ) +
  labs(
    x = "Model",
    y = "Estimated population size"
  )

grid.arrange(mark_plot, JAGS_plot)

```

```{r Table and AICc}
library(dplyr)

mark <- mark_estimates %>%
  mutate(MARK = sprintf("%.1f (%.1f–%.1f)", N_Estimate, N_Lower, N_Upper)) %>%
  select(Model, MARK, AICc)

jags <- JAGS_estimates %>%
  mutate(JAGS = sprintf("%.1f (%.1f–%.1f)", N_Estimate, N_Lower, N_Upper)) %>%
  select(Model, JAGS)

results_table <- full_join(mark, jags, by = "Model") %>%
  select(Model, MARK, JAGS, AICc)

results_table

```

The time-varying capture probability model ($M_t$) has the lowest AICc based on MARK outputs.
